Regionally Additive Models: Explainable-by-design models minimizing feature interactions

### 2023-06-29 report/main_paper.tex
latest paper

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{geometry}[margin=1in]
\usepackage{subcaption}
\usepackage{array}
\usepackage{multirow}
\usepackage{hyperref}

@@

\usepackage{natbib}
\usepackage{multicol}
\usepackage{algpseudocode}
\bibliographystyle{plainnat}
% set path for bibliography

\title{Regionally Additive Models: Explainable-by-design models minimizing feature interactions}

@@

\author{Vasilis Gkolemis}

\begin{document}
    \def\MakeUppercaseUnsupportedInPdfStrings{\scshape}
\maketitle

\begin{abstract}
Generalized Additive Models (GAMs) are widely utilized explainable-by-design models in various applications.
However, their assumption of independence among features can lead to suboptimal performance when violated.
To overcome this limitation, we introduce Regionally Additive Models (RAMs), a novel class of explainable-by-design models.
RAMs mitigate the issue by identifying subregions in the feature space where interactions are minimized and fitting multiple GAMs accordingly.
The RAM framework consists of three steps.
Firstly, we train a black-box model.
Secondly, using Regional Effect Plots, we identify subregions where the black-box model exhibits near-local additivity.
In these subregions, the effect of each feature on the target is independent of the values of other features.
Lastly, we fit a GAM specifically for each identified subregion.
We validate the effectiveness of RAMs through experiments on both a synthetic example and real-world datasets.
The results confirm that RAMs offer improved expressiveness compared to GAMs while maintaining interpretability.
\end{abstract}

\section{Introduction}

@@

\begin{algorithm}
\caption{DetectSubregions}
\label{alg:subregion_detection}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\BlankLine
\Input{Dataset $X$, Gradients $J$, Maximum depth $T$, Feature $s$}
\BlankLine
\Output{Subregions $\{\mathcal{R}_{st}\}_{t=1}^{T_s}$, where $T_s \leq 2^T$}
\BlankLine
$\mathcal{H}_s^0$ \tcp*{Compute the level of interactions before any split}
$T_s = 0$ \tcp*{Initialize the number of splits for feature $s$}
\For{$l = 1$ to $L$}{
    \If{H_{s}^{l-1} = 0}{
        break\;
    }
    \tcc{Find best split feature $c_s^l$ at point $p_s^l$, leading to loss
        $\mathcal{H}_s^l$ using regions of previous level}
    Find $\mathcal{H}_{st}, c, p$ of the optimal split based on $\mathcal{R}_s^l$ \;
    \If{$1 - \frac{\mathcal{H}_s^l}{\mathcal{H}_s^{l-1}} > \epsilon$}{
        break\;
    }
    $T_s = 2^l$ \tcp*{Update the number of splits for feature $s$}
}
\Return{$\left\{\mathcal{R}_{st}|s\in\left\{1,\ldots,D\right\},
  t\in\left\{1,\ldots,T_s\right\}\right\}$}
\end{algorithm}

%\begin{algorithm}
%\caption{BestSplit}
%\label{alg:single_feature_subregion}
%\SetAlgoLined
%\SetKwInOut{Input}{Input}
%\SetKwInOut{Output}{Output}
%
%\Input{X\_list, J\_list, s, c, is\_cat}
%\Output{BestSplits}
%
%% Initialize variables
%\BlankLine
%L[c, p] $\gets$ None\; % \Comment{List to store interaction values}
%\For{c $s = 1$ to $D$ \textnormal{if} c $\neq$ s}{
%    is\_cat $\gets$ IsCategorical(X\_list, c)\;
%    positions $\gets$ GetPositions(X\_list, c, is\_cat)\;
%    L\_pos $\gets$ []\; % \Comment{List to store interaction values}
%    \For{p \textnormal{in} positions}{
%        X\_split, J\_split $\gets$ SplitDataset(X\_list, J\_list, c, p, is\_cat)\;
%        L $\gets$ GetInteraction(X\_split, J\_split)\;
%    }
%}
%L\_min $\gets$ min(L)\;
%c\_min, p\_min = argmin(L)\;
%is\_cat $\gets$ IsCategorical(X\_list, c\_min)\;
%X\_split, J\_split $\gets$ SplitDataset(X\_list, J\_list, c\_min, p\_min, is\_cat)\;
%
%\Return{L\_min, p\_min c\_min, X\_split, J\_split}
%\BlankLine
%\end{algorithm}
%
%\begin{algorithm}
%\caption{GetInteraction}
%\label{alg:get_interaction}
%\SetAlgoLined
%\SetKwInOut{Input}{Input}
%\SetKwInOut{Output}{Output}
%\Input{$X\_list$, $J\_list$, $min\_points$}
%\Output{weighted average of interaction levels}
%N $\gets$ total number of items in $X\_list$\;
%W $\gets$ []\;
%L $\gets$ []\;
%\For{$i$ in len($X\_list$)}{
%    X $\gets$ X\_list[i]\;
%    J $\gets$ J\_list[i]\;
%    \uIf{$|X|$ < min\_points}{
%      Append $\infty$ to L\;
%    }
%    \Else{
%        Append $\mathcal{L}(X, J)$ to $L$\;
%    }
%    Append $\frac{|X|}{N}$ to $W$\;
%}
%\Return $\text{sum}(L \cdot W)$\;
%\end{algorithm}

@@

We evaluate the proposed approach on two typical tabular datasets, namely the Bike-Sharing Dataset~\citep{misc_bike_sharing_dataset_275}
and the California Housing Dataset~\citep{pace1997sparse}.

\begin{table}[htbp]
  \centering
  \caption{The table compares RMSE values of different models on two datasets: Bike-Sharing and California Housing.
  The models include DNN, GAM, RAM, GAM², and RAM² (representing 2nd order interactions).
  Lower values indicate better performance.
  RAM consistently outperforms GAM and approaches DNN performance.}
  \label{tab:sample}
  \begin{tabular}{l|c|cccc}
      \hline
      & \textbf{Black-box} & \multicolumn{4}{c}{\textbf{x-by-design}} \\
      \hline
      \hline
      & all orders & \multicolumn{2}{c}{1\textsuperscript{st} order} & \multicolumn{2}{c}{2\textsuperscript{nd} order} \\
      \hline
      \hline
      & \textbf{DNN} & \textbf{GAM} & \textbf{RAM} & \textbf{GAM}$^2$ & \textbf{RAM}$^2$ \\
      \hline
      Bike Sharing      & 0.254 & 0.549 & 0.430 & 0.298 & 0.278 \\
      \hline
      California Housing      & 0.369 & 0.600 & 0.535 & 0.554 & 0.514 \\
  \end{tabular}
\end{table}

@@

% Training - Evaluation - Region Extraction
As a black-box model, we train for \(60\) epochs a fully-connected Neural Network with 6 hidden layers, using the Adam optimizer with a learning rate of $0.001$.
The model attains a root mean squared error of \( 0.25 \cdot 181 \approx 45\) counts on the test set.
Subsequently, we extract the subregions, searching for splits up to a maximum spliting depth of \(T=3\).
Following the postprocessing step, we find that the only split that substantially reduces the level of interactions within the subregions is based on the feature
$X_{\mathtt{hour}}$. This feature is divided into two subgroups: $X_{\mathtt{hour}} | \when{X_{\mathtt{workingday}} \neq 1}$ and $X_{\mathtt{hour}} \when{X_{\mathtt{workingday} = 1}}$.

Figure~\ref{fig:bike_sharing} clearly illustrates that the impact of the hour of the day on bike rentals varies
significantly depending on whether it is a working day or a non-working day.
Specifically, during working days, there is higher demand for bike rentals in the morning and afternoon hours,
which aligns with the typical commuting times (\ref{subfig:bike_rentals_regional_gam_1}).
On the other hand, during non-working days, bike rentals peak in the afternoon as individuals engage in
leisure activities (\ref{subfig:bike_rentals_regional_gam_2}).
The proposed RAM method effectively captures and detects this interaction by establishing two distinct subregions,
each corresponding to working days and non-working days, respectively.
Subsequently, the EBM that is fitted to each subregion, successfully learns these patterns,
achieving a root mean squared error of approximately \( 0.43 \cdot 181 \approx 77\) counts on the test set.
It is noteworthy that RAM not only preserves the interpretability of the model,
but it also enhances the interpretation of the underlying modeling process.
By identifying and highlighting the interaction between the hour of the day and the day type,
RAM provides valuable insights into the relationship between these variables and their influence on bike rentals.
In contrast, the GAM model~\ref{subfig:bike_rentals_gam} is not able to capture this interaction and
achieves a root mean squared error of \( 0.55 \cdot 181 \approx 100\) counts on the test set.
Finally, in table~\ref{tab:sample}, we also observe that the RAM$^2$, i.e., RAM with second-order interactions,
outperforms the equivalent GAM$^2$ model in terms of predictive performance.

@@

As a black-box model, we train for $45$ epochs a fully-connected Neural Network with 6 hidden layers,
using the Adam optimizer with a learning rate of $0.001$.
The model achieves a root mean square error (RMSE) of about \(40\)K dollars on the test set.
Subsequently, we perform subregion extraction by searching for splits up to a maximum depth of \(T=3\).
After the postprocessing step, we discover that several splits significantly reduce the level of interactions,
resulting in an expanded input space consisting of \(16\) features, as we show in table~\ref{tab:california_housing_subregions}.
Out of them, we randomly select and illustrate in Figure~\ref{fig:california_housing} the effect of the feature $X_{\mathtt{long}}$.
As we observe, for the house blocks located in the southern part of California ($X_{\mathtt{lat}} \leq 34.9$),
the house value decreases in an almost linear fashion as we move eastward ($X_{\mathtt{long}}$ increases).
In contrast, for the house blocks located in the northern part of California ($X_{\mathtt{lat}} > 34.9$),
the house value performs a rapid (non-linear) decrease as we move eastward ($X_{\mathtt{long}}$ increases).
We also observe that although the EBM fitted to each subregion captures the general trend,
it does not align perfectly with the regional effect.
As in the Bike-Sharing Example, the RMSE of the RAM model, i.e. \( 0.53 \cdot 110 \cdot 10^3 \approx 58.3 \cdot 10^3\) dollars on the test set,
is lower than the one of the GAM model, i.e.\( 0.6 \cdot 110 \cdot 10^3 \approx 66 \cdot 10^3\) dollars.
These results indicate that the RAM model provides superior predictions compared to the GAM model.
The same conclusion holds is when comparing the RAM$^2$ and the GAM$^2$ models.
\begin{table}[htbp]
  \centering
  \caption{California Housing: Subregions Detected by RAM}
  \label{tab:california_housing_subregions}
  \begin{tabular}{c|c}
      Feature & Subregions \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{long}}$} & $X_{\mathtt{long}} \when{X_{\mathtt{lat}} \leq 34.9}$ \\
      & $X_{\mathtt{long}} \when{X_{\mathtt{lat}} > 34.9}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{lat}}$} & $X_{\mathtt{lat}} \when{X_{\mathtt{long}} \leq -120.31}$ \\
      & $X_{\mathtt{lat}} \when{X_{\mathtt{long}} > -120.31}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{total\_rooms}}$} & $X_{\mathtt{total\_rooms}} \when{X_{\mathtt{total\_bedrooms}} \leq 449.37}$ \\
        & $X_{\mathtt{total\_rooms}} \when{X_{\mathtt{total\_bedrooms}} > 449.37}$ \\
      \hline
      \multirow{4}{*}{$X_{\mathtt{total\_bedrooms}}$} & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} \leq 411} \when{X_{\mathtt{total\_bedrooms}} \leq 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} \leq 411} \when{X_{\mathtt{total\_bedrooms}} > 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} > 411} \when{X_{\mathtt{total\_bedrooms}} \leq 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} > 411} \when{X_{\mathtt{total\_bedrooms}} > 647}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{population}}$} & $X_{\mathtt{population}} \when{X_{\mathtt{households}} \leq 411.5}$ \\
      & $X_{\mathtt{population}} \when{X_{\mathtt{households}} > 411.5}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{households}}$} & $X_{\mathtt{households}} \when{X_{\mathtt{total\_bedrooms}} \leq 630.57}$ \\
        & $X_{\mathtt{households}} \when{X_{\mathtt{total\_bedrooms}} > 630.57}$ \\
  \end{tabular}
\end{table}

@@

Another important area of investigation involves exploring the impact of second-order effects within the RAM framework.
While our experimenation demonstrated that even with the current subregion detection, RAM$^2$s outperform GAM$^2$s,
it may be the case, that for second-order models the optimal subregions are not necessarily those that maximize the additive effect of individual features,
but rather those that maximize the additive effect of feature pairs.

%\section{Appendix}
%
%\subsection{Regional DALE formulation}
%\label{sec:regional_dale}
%
%Below we provide the regional DALE formulation for the bin-effect, the bin-deviation and the aggregated deviation on a subregion \(\mathcal{R}_{st}\):
%
%\begin{equation}
%    \label{eq:bin_effect_subregion}
%    \hat{\mu}_{\mathcal{R}_{st}}(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k \cap \mathcal{R}_{st}|} \sum_{i:\mathbf{x}^i \in \mathcal{S}_k \cap \mathcal{R}_{st} } \dfdx(\mathbf{x}^i)
%\end{equation}
%
%\begin{equation}
%  \label{eq:bin_deviation_subregion}
%  \hat{\sigma}_{\mathcal{R}_{st}}^2(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k \cap \mathcal{R}_{st}| - 1}
%\sum_{i:\mathbf{x}^i \in \mathcal{S}_k \cap \mathcal{R}_{st} } \left ( \dfdx(\mathbf{x}^i) -
%    \hat{\mu}(z_{k-1}, z_k) \right )^2
%\end{equation}
%
%\begin{equation}
%  \label{eq:SE}
%  \mathcal{H}_{\mathcal{R}_{sk}} = \sum_{k=1}^K (z_k - z_{k-1})^2 \hat{\sigma}_{\mathcal{R}_{st}}^2(z_{k-1}, z_k)
%\end{equation}

%\subsection{Algorithmic Details of Subregion Detection}
%
%\begin{algorithm}
%\caption{SplitDataset}
%\label{alg:split_dataset}
%\SetAlgoLined
%\SetKwInOut{Input}{Input}
%\SetKwInOut{Output}{Output}
%\Input{X\_list, J\_list, c, val, is\_cat}
%\Output{X\_split, J\_split}
%X\_split $\gets$ []\;
%J\_split $\gets$ []\;
%\For{i = 1 to len(X\_list)}{
%    X= X\_list[i]\;
%    J= J\_list[i]\;
%    \uIf{is\_cat}{
%        ind\_1 $\gets$ X[:, c] $=$ val\;
%        ind\_2 $\gets$ X[:, c] $\neq$ val\;
%    }
%    \Else{
%        ind\_1 $\gets$ X[:, c] $\leq$ val\;
%        ind\_2 $\gets$ X[:, c] $>$ val\;
%    }
%    Append X[ind\_1], X[ind\_2] to X\_split\;
%    Append J[ind\_1], J[ind\_2] to J\_split\;
%}
%\Return X\_split, J\_split
%\end{algorithm}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{geometry}[margin=1in]
\usepackage{subcaption}
\usepackage{array}
\usepackage{multirow}
\usepackage{hyperref}

@@

\usepackage{natbib}
\usepackage{algpseudocode}
\bibliographystyle{plainnat}
% set path for bibliography

@@

\begin{document}
%
\title{Regionally Additive Models: Explainable-by-design models minimizing feature interactions}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
%    First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%Third Author\inst{3}\orcidID{2222--3333-4444-5555}
}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
%    Princeton University, Princeton NJ 08544, USA \and
%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{lncs@springer.com}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Generalized Additive Models (GAMs) are widely used explainable-by-design models in various applications.
GAMs assume that the output can be represented as a sum of univariate functions, referred to as components.
However, this assumption fails in ML problems where the output depends on multiple features simultaneously.
In these cases, GAMs fail to capture the interaction terms of the underlying function, leading to subpar accuracy.
To (partially) address this issue, we propose Regionally Additive Models (RAMs), a novel class of explainable-by-design models.
RAMs identify subregions within the feature space where interactions are minimized.
Within these regions, it is more accurate to express the output as a sum of univariate functions (components).
Consequently, RAMs fit one component per subregion of each feature instead of one component per feature.
This approach yields a more expressive model compared to GAMs while retaining interpretability.
The RAM framework consists of three steps.
Firstly, we train a black-box model.
Secondly, using Regional Effect Plots, we identify subregions where the black-box model exhibits near-local additivity.
Lastly, we fit a GAM component for each identified subregion.
We validate the effectiveness of RAMs through experiments on both synthetic and real-world datasets.
The results confirm that RAMs offer improved expressiveness compared to GAMs while maintaining interpretability.
\keywords{Explainable AI \and Generalized Additive models \and x-by-design}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:intro}

@@

\begin{algorithm}
\caption{DetectSubregions}
\label{alg:subregion_detection}
\SetAlgoLined
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\BlankLine
\Input{Dataset $X$, Gradients $J$, Maximum depth $T$, Feature $s$}
\BlankLine
\Output{Subregions $\{\mathcal{R}_{st}\}_{t=1}^{T_s}$, where $T_s \leq 2^T$}
\BlankLine
$\mathcal{H}_s^0$ \tcp*{Compute the level of interactions before any split}
$T_s = 0$ \tcp*{Initialize the number of splits for feature $s$}
\For{$l = 1$ to $L$}{
    \If{$H_{s}^{l-1} = 0$}{
        break\;
    }
    \tcc{Find best split feature $c_s^l$ at point $p_s^l$, leading to loss
        $\mathcal{H}_s^l$ using regions of previous level}
    Find $\mathcal{H}_{st}, c, p$ of the optimal split based on $\mathcal{R}_s^l$ \;
    \If{$1 - \frac{\mathcal{H}_s^l}{\mathcal{H}_s^{l-1}} > \epsilon$}{
        break\;
    }
    $T_s = 2^l$ \tcp*{Update the number of splits for feature $s$}
}
\Return{$\left\{\mathcal{R}_{st}|s\in\left\{1,\ldots,D\right\},
  t\in\left\{1,\ldots,T_s\right\}\right\}$}
\end{algorithm}

@@

We evaluate the proposed approach on two typical tabular datasets:
the Bike-Sharing Dataset~\citep{misc_bike_sharing_dataset_275} and the California Housing Dataset~\citep{pace1997sparse}.

\begin{table}[htbp]
  \centering
  \caption{The table compares the Mean Absolute Error (MAE) and the Root Mean Square Error (RMSE) of
  DNN, GAM, RAM, GA$^2$M, and RA$^2$M (representing 2nd order interactions), on two datasets: Bike-Sharing and California Housing.
  Lower values indicate better performance.
  RAM consistently outperforms GAM and approaches DNN performance.}
  \label{tab:sample}
  \begin{tabular}{l|c|cccc}
      & \textbf{Black-box} & \multicolumn{4}{c}{\textbf{x-by-design}} \\
      \hline
      \hline
      & all orders & \multicolumn{2}{c}{1\textsuperscript{st} order} & \multicolumn{2}{c}{2\textsuperscript{nd} order} \\
      \hline
      \hline
      & \textbf{DNN} & \textbf{GAM} & \textbf{RAM} & \textbf{GA}$^2$\textbf{M} & \textbf{RA}$^2$\textbf{M} \\
      \hline
      Bike Sharing (MAE)  & 0.254 & 0.549 & 0.430 & 0.298 & 0.278 \\
      Bike Sharing (RMSE) & 0.389 & 0.734 & 0.563 & 0.438 & 0.412 \\
      \hline
      California Housing (MAE)  & 0.373 & 0.600 & 0.553 & 0.554 & 0.533 \\
      California Housing (RMSE) & 0.533 & 0.819 & 0.754 & 0.774 & 0.739 \\
  \end{tabular}
\end{table}

@@

% Training - Evaluation - Region Extraction
As a black-box model, we train for \(60\) epochs a fully-connected Neural Network with 6 hidden layers, using the Adam optimizer with a learning rate of $0.001$.
The model attains a root mean squared error of \( 0.39 \cdot 181 \approx 70\) counts on the test set.
Subsequently, we extract the subregions, searching for splits up to a maximum spliting depth of \(T=3\).
Following the postprocessing step, we find that the only split that substantially reduces the level of interactions within the subregions is based on the feature
$X_{\mathtt{hour}}$. This feature is divided into two subgroups: $X_{\mathtt{hour}} | \when{X_{\mathtt{workingday}} \neq 1}$ and $X_{\mathtt{hour}} \when{X_{\mathtt{workingday} = 1}}$.

Figure~\ref{fig:bike_sharing} clearly illustrates that the impact of the hour of the day on bike rentals varies
significantly depending on whether it is a working day or a non-working day.
Specifically, during working days, there is higher demand for bike rentals in the morning and afternoon hours,
which aligns with the typical commuting times (Figure~\ref{subfig:bike_rentals_regional_gam_1}).
On the other hand, during non-working days, bike rentals peak in the afternoon as individuals engage in
leisure activities (Figure~\ref{subfig:bike_rentals_regional_gam_2}).
The proposed RAM method effectively captures and detects this interaction by establishing two distinct subregions,
each corresponding to working days and non-working days, respectively.
Subsequently, the EBM that is fitted to each subregion, successfully learns these patterns,
achieving a root mean squared error of approximately \( 0.56 \cdot 181 \approx 101\) counts on the test set.
It is noteworthy that RAM not only preserves the interpretability of the model,
but it also enhances the interpretation of the underlying modeling process.
By identifying and highlighting the interaction between the hour of the day and the day type,
RAM provides valuable insights into the relationship between these variables and their influence on bike rentals.
In contrast, the GAM model~\ref{subfig:bike_rentals_gam} is not able to capture this interaction and
achieves a root mean squared error of \( 0.73 \cdot 181 \approx 132\) counts on the test set.
Finally, in table~\ref{tab:sample}, we also observe that the RA$^2$M, i.e., RAM with second-order interactions,
outperforms the equivalent GA$^2$M model in terms of predictive performance.
Specifically, the RA$^2$M model achieves a root mean squared error of \( 0.41 \cdot 181 \approx 74\) counts,
while the GA$^2$M model of \( 0.44 \cdot 181 \approx 80\) counts on the test set.
It is worth noticing that the RA$^2$M model's accuracy is comparable to the black-box model's accuracy.

@@

As a black-box model, we train for $45$ epochs a fully-connected Neural Network with 6 hidden layers,
using the Adam optimizer with a learning rate of $0.001$.
The model achieves a root mean square error (RMSE) of about \(58\)K dollars on the test set.
Subsequently, we perform subregion extraction by searching for splits up to a maximum depth of \(T=3\).
After the postprocessing step, we discover that several splits significantly reduce the level of interactions,
resulting in an expanded input space consisting of \(16\) features, as we show in table~\ref{tab:california_housing_subregions}.
Out of them, we randomly select and illustrate in Figure~\ref{fig:california_housing} the effect of the feature $X_{\mathtt{long}}$.
As we observe, for the house blocks located in the southern part of California ($X_{\mathtt{lat}} \leq 34.9$),
the house value decreases in an almost linear fashion as we move eastward ($X_{\mathtt{long}}$ increases).
In contrast, for the house blocks located in the northern part of California ($X_{\mathtt{lat}} > 34.9$),
the house value performs a rapid (non-linear) decrease as we move eastward ($X_{\mathtt{long}}$ increases).
We also observe that although the EBM fitted to each subregion captures the general trend,
it does not align perfectly with the regional effect.
As in the Bike-Sharing Example, the RMSE of the RAM model, i.e. \( 0.75 \cdot 110 \approx 82.5\)K dollars on the test set,
is lower than the one of the GAM model, i.e.\( 0.82 \cdot 110 \approx 90\)K dollars.
These results indicate that the RAM model provides superior predictions compared to the GAM model.
The same conclusion holds is when comparing the RA$^2$M and the GA$^2$M models,
where the former achieves a RMSE of \( 0.74 \cdot 110 \approx 81\)K dollars,
while the latter of \( 0.77 \cdot 110 \approx 85\)K dollars.

\begin{table}[htbp]
  \centering
  \caption{California Housing: Subregions Detected by RAM}
  \label{tab:california_housing_subregions}
  \begin{tabular}{c|c}
      Feature & Subregions \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{long}}$} & $X_{\mathtt{long}} \when{X_{\mathtt{lat}} \leq 34.9}$ \\
      & $X_{\mathtt{long}} \when{X_{\mathtt{lat}} > 34.9}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{lat}}$} & $X_{\mathtt{lat}} \when{X_{\mathtt{long}} \leq -120.31}$ \\
      & $X_{\mathtt{lat}} \when{X_{\mathtt{long}} > -120.31}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{total\_rooms}}$} & $X_{\mathtt{total\_rooms}} \when{X_{\mathtt{total\_bedrooms}} \leq 449.37}$ \\
        & $X_{\mathtt{total\_rooms}} \when{X_{\mathtt{total\_bedrooms}} > 449.37}$ \\
      \hline
      \multirow{4}{*}{$X_{\mathtt{total\_bedrooms}}$} & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} \leq 411} \when{X_{\mathtt{total\_bedrooms}} \leq 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} \leq 411} \when{X_{\mathtt{total\_bedrooms}} > 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} > 411} \when{X_{\mathtt{total\_bedrooms}} \leq 647}$ \\
        & $X_{\mathtt{total\_bedrooms}} \when{X_{\mathtt{households}} > 411} \when{X_{\mathtt{total\_bedrooms}} > 647}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{population}}$} & $X_{\mathtt{population}} \when{X_{\mathtt{households}} \leq 411.5}$ \\
      & $X_{\mathtt{population}} \when{X_{\mathtt{households}} > 411.5}$ \\
      \hline
      \multirow{2}{*}{$X_{\mathtt{households}}$} & $X_{\mathtt{households}} \when{X_{\mathtt{total\_bedrooms}} \leq 630.57}$ \\
        & $X_{\mathtt{households}} \when{X_{\mathtt{total\_bedrooms}} > 630.57}$ \\
  \end{tabular}
\end{table}

@@

Another important area of investigation involves exploring the impact of second-order effects within the RAM framework.
While our experimenation demonstrated that even with the current subregion detection, RA$^2$Ms outperform GA$^2$Ms,
it may be the case, that for second-order models the optimal subregions are not necessarily those that maximize the additive effect of individual features,
but rather those that maximize the additive effect of feature pairs.

\subsubsection{Acknowledgements} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).
